---
title: "Class 6.1 Example in class - Linear modeling dos and don'ts"
output: html_notebook
---

Loading data
```{r}
library(statisticalModeling)
library(mosaicModel)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(broom)
library(lattice)
ads_sales <- as_tibble(read.csv("Advertising.csv"))
ads_sales
```

## Oversimplified Modeling process

1. Fit a model
2. Evaluate the model by adding residuals and looking at MSE
3. Use model for decision


# Less naive modeling
1. Exploratory data analysis

*Split data into training and test

2. Fit a model
3. Evaluate the model by adding residuals and looking at MSE
  *Using cross-validation
  
Interpret model
  *Using bootstrapped confidence intervals for effect sizes / parameters
  
4. Use model for decision

## Exploratory Data Analysis
Before doing any modeling, you should get a basic feel for dependent and indpendent variables and their relationships visually.
```{r}
#look at the data with histograms and scatterplots

car::scatterplotMatrix(ads_sales)


```
Write code for each of the below:
```{r}

# What is the number of males/females in the dataset? What does this already tell you...?

# What is the mean salary by sex? Hint: you'll have to groupby sex (`sx`)

# Draw histograms for the distribution of salaries for males and females (separately)
# Hint: you can use ggplot and facet 
# The x and y axes should be consistent between the graphs

# Draw histograms for the distribution of salaries by rank

# Create scatterplots to show how salary compares to years since degree / in current rank

```
##Simple linear regression: what is the salary increase associated with each additional year in your current position (yr)?

```{r}
# Create a simple linear model that assesses the relationship tv radio, and newspapers and sales



# What is the effect size of years on salary?
# Hint: you can use tools from statisticalModeling package used in the Datacamp course

# Evaluate the accuracy of your model. Calculate a metric for it.

# Interpret the metric for accuracy above.

# Draw a scatterplot with your model on it to show how well the model fits the data
# Hint: broom and ggplot may be useful here
```
## Skeptical Modeling process
1. Exploratory data analysis

Split data into training and test

2. Fit a model
3. Evaluate the model by adding residuals and looking at MSE
  Using cross-validation
  
Interpret model
  Using bootstrapped confidence intervals for effect sizes / parameters
  
4. Use model for decision


```{r}
# Using multiple regression, create a linear model that uses 
# sex, rank, and years in current rank variables to estimate salary

# What is the effect size for each variable in the model?

# Write what surprises you....
# (write text here)

# Create a new dataset with your multiple regression model's predictions in a new column


# How do the new model's predictions compare to the simpler model's predictions?
# Make plots and then write below

# (write text here)


# Create a scatter plot with years since degree on the x axis, and 
# Salary on the y axis. Show points for both the data, and the multiple regression values.
# Bonus points: improve the visualization above by adding more variables
```
##Assess predictions

```{r}
# Make a scatterplot that compares the data (x) to the predictions (y)
# Add a line showing where the perfect values would be (i.e., prediction equal to data)

# What are the MSE values of the two models?

# Which one explains more variance?

# Plot the salary v.s. the *residuals* (difference between actual and predicted values)
# Create a separate (adjacent) plot for each model
# Add a horizontal line at 0 to help interpret the graph
# or use the ggplot resources in #help on slack to make a better residuals plot
# ie
# http://shinyapps.org/apps/RGraphCompendium/index.php
# http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html

# Provide an interpretation for these graphs
# (write here)
```

## Oversimplified Modeling process
1. Exploratory data analysis
2. Fit a model
3. Evaluate the model by adding residuals and looking at MSE
4. Use model for decision


# Class Example

# Oversimplified Modeling - Example of what not to do


# Fit a model

```{r}
#fit a linear model for sales


# get the coefficients from the model
# tidy is a trick to make the model object from lm into a data frame
#   makes it easier to extract parts of it
#   -> you can access them with paramters[1] to get the intercept, [2] for the slope



```
# Evaluate the model

```{r}


```

## Interpret the model
```{r}


```

Do you think this is a good model for the data set? Why?
  Write your answer here

What information would help you answer the question above better?
  Write your answer here
  